---
title: "Analysis of beer data frame"
output: html_notebook
---

**Beer** a dataset with the brand, brewer, alcohol percentage, calories and grams of carbohydrates for 172 beers.

Questions - do brewers specialise in high alcohol beers? What are the higest in calorie / carbohydrate

```{r}
# Load the libraries that might be needed---------------------------------------

library(CodeClanData)
library(tidyverse)
library(styler)
library(psych)
```

```{r}
# Look at the structure of the beer data frame----------------------------------

str(beer)
```

The field `calories` is a `chr` type but looks to contain numbers - will change to numeric.
```{r}
# Changing field to numeric ----------------------------------------------------

beer$calories <- as.numeric(beer$calories)
```

```{r}
# Look at the summary to see the range of the numeric fields--------------------

summary(beer)
```

The mean is a little higher that the median for percent of alcohol and calories 
but for carbohydrates is almost eactly the same.

```{r}
quantile(beer$percent, seq(from = 0, to = 1, by = 0.20))
```
```{r}
quantile(beer$carbohydrates, seq(from = 0, to = 1, by = 0.20))
```

```{r}
quantile(beer$calories, seq(from = 0, to = 1, by = 0.20))
```


```{r}
# Use a histogram plot to see the shape of the numeric distributions------------

hist(beer$percent,
  main = "Histogram of beer alcohol percentage",
  xlab = "Pct (alcohol)"
)
hist(beer$carbohydrates,
  main = "Histogram of beer carbohydrates",
  xlab = "Carbohydrates (grams)"
)
hist(beer$calories,
  main = "Histogram of beer calories",
  xlab = "Calories"
)
```

Almost 60% percent of the beers in the dataset are in the 5% alcohol range.

Almost 47% percent of the beers in the dataset are in 11 - 15 gram of 
carbohydrates range. The histogram is slightly right skewed as there are small 
number of beers in the dataset that have very high carbohydrate values of almost
double that.

The calories histogram is also right skewed with the majority of values falling
in the range of just over 100 to 200 calories. There are a small number of beers
with values over 300.


```{r}
# Looking at the entries at the upper and lower ends of the histogram
# for percentage alcohol--------------------------------------------------------

beer %>%
  filter(percent > 10 | percent < 2)
```
```{r}
# Looking at the entries at the upper and lower end of the histogram for
# carbohydrates in beer---------------------------------------------------------

beer %>%
  filter(carbohydrates > 25 | carbohydrates < 2)
```
```{r}
# Looking at the entries at the upper end of the histogram for calories---------

beer %>%
  filter(calories > 250)
```
Flying Dog Brewery appear in the upper end of both calories and alcohol
percentage while Sierra Nevada appear at the top for calories.


What relationship do the numeric variables have?
```{r}
# Using scatterplots to visualise the relationships among numeric fields in the
# dataset. The intersection of each row and column holds the scatterplot of the
# variables indicated by the row and column pair ------------------------------

pairs(beer[c("calories", "percent", "carbohydrates")])
```


```{r}
# Producing a correlation matrix to try and see relationships more clearly.
# The diagonal now contains histograms depicting the distribution of values.
# The oval-shaped object in each scatterplot is a correlation ellipse. It
# provides a visualisation of correlation strength. The more the elipse is
# stretched, the stronger the correlation --------------------------------------

pairs.panels(beer[c("calories", "percent", "carbohydrates")])
```
In the above plot, the strongest correlation appears to be between percentage of 
alcohol and number of calories.



```{r}
# A faceted scatter plot to show all the brewers in the dataset and the
# alcohol percentage and calories variables ------------------------------------

ggplot(data = beer) +
  geom_point(mapping = aes(x = calories, y = percent)) +
  facet_wrap(~brewer)
```
Flying Dog Brewery seem to be focusing on a spread of products going all the 
way up to higher alcohol / higher calorie beers.

Also in the plot you can see a number of brewers that are focussed on a specific 
product (only one point showing on the plot in those cases).





















































