---
title: "Analysis of Game of Thrones list"
output: html_notebook
---

*Game of Thrones* Data on the book "A Game of Thrones", including names and 
genders for all of the characters.

Questions - 
Can we look at the balance of genders?
Are there any duplicates in the list of characters?

```{r}
# Load the libraries that might be needed---------------------------------------

library(CodeClanData)
library(tidyverse)
library(styler)
library(tidyr)
library(tidytext)
```


```{r}
# Looking at a summary of the data first ---------------------------------------

summary(game_of_thrones)
```


```{r}
# Just selecting the character part of the list --------------------------------

game_of_thrones[["characters"]]
```

```{r}
# Turn the characters section of the list into a data frame --------------------

got_characters <- reduce(game_of_thrones[["characters"]], bind_rows)
```

```{r}
# Plot the numebr of male and female characters --------------------------------

ggplot(data = got_characters) +
  stat_count(mapping = aes(x = gender))
```
In the data provided, it would appear there's a much greater number of male 
characters than female characters.


```{r}
# Looking for characters who appear in the list more than once -----------------

table(got_characters$name)
```


```{r}
# Filtering character names that appear in the list more than once -------------

got_duplicates <- got_characters %>%
  group_by(name) %>%
  filter((n() > 1))
got_duplicates
```

```{r}
# Using the above filtered data to create a plot by character name for those
# that appear in the list more than once ---------------------------------------

ggplot(data = got_duplicates) +
  stat_count(mapping = aes(x = name))
```
Is this data correct or are there duplicates?

```{r}
extract(got_characters, name, c("first_name", "last_name"), "([^ ]+) (.*)")
```

```{r}
# Trying to split out first_name and last_name ---------------------------------

got_characters_split <- extract(
  got_characters, name, c("first_name", "last_name"), "([^ ]+) (.*)"
)
```

```{r}
# Removing first name

got_characters_split$first_name <- NULL
```

```{r}
got_family_names <- got_characters_split %>%
  filter(!is.na(last_name) & last_name != "I") %>%
  group_by(last_name) %>%
  count(last_name, sort = TRUE)
got_family_names
```

I assume these must be important families in the Game of Thrones story as they
seem to occur the most often.


