---
title: "Analysis of temperature matrix"
output: html_notebook
---

*Temp* A matrix which contains the maximum temperature for each month in Scotland over 106 years (1910 to 2015). Each row corresponds to a year and each column to a month (January to December).

```{r}
# Looking at the data ----------------------------------------------------------

head(temp)
```

```{r}
# Creating a summary of the data -----------------------------------------------
summary(temp)
```



```{r}
# Creating the columns to add to the matrix-------------------------------------

# Calulate the values first before joining to the matrix------------------------

total_of_max_temps <- rowSums(temp)
avg_of_max_temps <- rowMeans(temp)
year_of_measurement <- seq(from = 1910, to = 2015, by = 1)
```

```{r}
# Adding year to matrix----------------------------------------------------------

max_temps_scotland <- cbind(year_of_measurement, temp)
```

```{r}
# Adding totals-----------------------------------------------------------------

max_temps_scotland <- cbind(total_of_max_temps, max_temps_scotland)
```

```{r}
# Adding averages---------------------------------------------------------------

max_temps_scotland <- cbind(avg_of_max_temps, max_temps_scotland)
```



```{r}
# Plot of max temperatures in Scotland------------------------------------------

ggplot(
  data = as.data.frame(
    max_temps_scotland
  ),
  mapping = aes(
    x = year_of_measurement,
    y = avg_of_max_temps
  )
) +
  geom_point() +
  geom_smooth()
```

Interestingly, this shows there was a peak at 1949 of 11.34 and then another 
peak in 1959 of 11.45. Average max temperatures didn't exceed the 1959 figure 
until 2003.

The geom_smooth uses a smooth line fitted to the data to show the trend - which 
seems to be steadily increasing since the mid 1970s.

It might be useful to try and show the monthly temparature changes each year 
using an animation with circles to represent temperature and and have the 
circles coloured to represent each month. The animation could play across the 
years so you could see the spread of temperatures moving in each year and also 
how each month was affected.


```{r}
# See if we can select summer data from just June, July and August for each year

head(temp[, 6:8])
```

```{r}
# Run this first to select the data for the matrix------------------------------

summer_temps_scotland <- temp[, 6:8]
```

```{r}
# Run this next so we are only calculating on the temperatures-----------------

total_of_max_summer_temps <- rowSums(summer_temps_scotland)
avg_of_max_summer_temps <- rowMeans(summer_temps_scotland)
```

```{r}
# Now we can add the years so they won't be in the calculation -----------------

summer_temps_scotland <- cbind(year_of_measurement, summer_temps_scotland)
```

```{r}
# Add the totals vector --------------------------------------------------------

summer_temps_scotland <- cbind(total_of_max_summer_temps, summer_temps_scotland)
```

```{r}
# Add the means vector ---------------------------------------------------------

summer_temps_scotland <- cbind(avg_of_max_summer_temps, summer_temps_scotland)
```


```{r}
# Plot of summer months June, July and August ----------------------------------

ggplot(
  data = as.data.frame(
    summer_temps_scotland
  ),
  mapping = aes(
    x = year_of_measurement,
    y = avg_of_max_summer_temps
  )
) +
  geom_point() +
  geom_smooth()
```

The trend looks a little flatter for summer months when comparing to the plot 
for the whole year. The plot also shows several outliers with high temperatures 
right from the 1930s onwards. These high temperatures appear to be larger 
outliers than the spread of the lower average maximum summer temperatures at the 
lower part of the plot.

1976 seems to be the year that everyone remembers as the long hot summer. 
However, for Scotland, 1995 was on overage slightly hotter for the summer 
months of June, July and August.

